---
title: "Machine Learning Project"
author: "Sofian Hamiti"
output: html_document
---

#Summary
The objective of the project is to provide 20 submissions that predict the exercises being performed in the test set.

We use the Random Forests method to fit our model.

The data for this project come from the paper:
Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.


###Loading libraries and enabling multi-core processing
```{r,message=FALSE,warning=FALSE}
library(caret)
library(randomForest)
library(doParallel)
set.seed(1337)
cl <- makeCluster(detectCores())
registerDoParallel(cl)
```


###Downloading and reading data 
```{r}
training.file <- "pml-training.csv"
test.file     <- "pml-testing.csv"
training.url  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test.url      <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(training.url, training.file, method="auto")
download.file(test.url, test.file, method="auto")
training       <- read.csv(training.file, na.strings = c("", "NA", "#DIV/0!") )
test           <- read.csv(test.file, na.strings = c("", "NA", "#DIV/0!") )
training       <- training[,-c(1,5,6)]
test           <- test[,-c(1,5,6)]
```

###Tidying data
We remove NA values on both data frames.
```{r}
training <- training[,sapply(training, function(x) !any(is.na(x)))]
training <- training[ , colSums(is.na(training)) < nrow(training)]
test <- test[,sapply(test, function(x) !any(is.na(x)))]
test <- test[ , colSums(is.na(test)) < nrow(test)]
```

###Partition training data into training.train and training.test sets (50% in size)
```{r}
partition  <- createDataPartition(training$classe, p=.50, list=FALSE)
training.train <- training[partition,]
training.test  <- training[-partition,]
```

###Fitting a Random Forests model
```{r}
model <- train(training.train[,-57],
                       training.train$classe,
                       tuneGrid=data.frame(mtry=3),
                       trControl=trainControl(method="none")
                       )
```

###Confusion Matrix for testing set
```{r}
confusionMatrix(predict(model,
                        newdata=training.test[,-57]),
                training.test$classe
                )
```
Accuracy: 0.9964, Kappa statistic: 0.9955.
We can consider the model as really good.

```{r}
plot(varImp(model))
```


```{r}
stopCluster(cl)
```


